---
title: "Benchmark"
author: Hung
date: "`r Sys.Date()`"
---

# Synopsis
This file performs the benchmarks

```{r}
# Install packages if needed
pkgs <- list(
  cran = c("missMDA", "softImpute", "missRanger"),
  bioc = c("methyLImp2", "impute"),
  github = c("hhp94/slideimp", "hhp94/slideimp.extra")
)

if (!rlang::is_installed("BiocManager")) install.packages("BiocManager")

for (pkg in pkgs$cran) {
  if (!rlang::is_installed(pkg)) install.packages(pkg)
}

for (pkg in pkgs$bioc) {
  if (!rlang::is_installed(pkg)) BiocManager::install(pkg)
}

for (pkg in pkgs$github) {
  pkg_name <- sub(".*/", "", pkg)
  if (!rlang::is_installed(pkg_name)) remotes::install_github(pkg)
}

# Load libraries
library(impute)
library(slideimp)
library(bench)
library(here)
library(qs2)
library(duckdb)
library(ggplot2)
library(tidyverse)
library(slideimp.extra)
source("utils.R")
```

# Multi Cores
* R2 requests that the benchmark done on multiple cores. We perform grouped imputation on the EPICv2 GSE286313 data using 4/8/12 cores in parallel.
* Scenarios will be K-NN, pca_imp, methyLImp2 PSOCK, and methyLImp2 FORK.
* We will also have to pin the jobs to a specific node for maximum reproducibility.

## Prep
```{r}
library(slideimp.extra)

beta_matrix <- do.call(
  rbind,
  lapply(
    get_DNAm(
      "EPICv2",
      # c(glue::glue("chr{1:8}"))
      c(glue::glue("all"))
    )$path,
    qs_read
  )
) |>
  t()

# beta_matrix <- dedup_matrix(beta_matrix, "EPICv2")
cv <- col_vars(beta_matrix)

beta_matrix |> dim()
beta_matrix <- beta_matrix[, !is.na(cv) & cv > 0]
beta_matrix |> dim()
beta_matrix <- beta_matrix[, (colSums(is.na(beta_matrix)) / nrow(beta_matrix)) < 0.9]
beta_matrix |> dim()

group_df <- group_features(beta_matrix, ilmn_manifest("EPICv2", dedupped = FALSE))
group_df <- dplyr::filter(group_df, !group %in% c("0", "M"))

qs2::qs_save(beta_matrix, here("GSE286313", "GSE286313.EPICv2.chr1-Y.qs2"))
qs2::qs_save(group_df, here("GSE286313", "GSE286313.EPICv2.chr1-Y.group_df.qs2"))
```

## Script
```{r benchmark.R, eval = F}
library(argparser)

p <- arg_parser("benchmark")
p <- add_argument(p, "--beta", help = "beta matrix file path", type = "character")
p <- add_argument(p, "--group", help = "group file path", type = "character")
p <- add_argument(p, "--cores", help = "number of cores", type = "integer")
p <- add_argument(
  p, "--method",
  help = "knn, knn_tree, knn_tree_chr, pca, missMDA, impute.knn, methyLImp2_fork, methyLImp2_psock",
  type = "character"
)
p <- add_argument(p, "--debug", help = "enable debug mode", flag = TRUE)
argv <- parse_args(
  p
  # ,
  # c(
  #   "--beta", here::here("GSE286313", "GSE286313.EPICv2.chr1-Y.qs2"),
  #   "--group", here::here("GSE286313", "GSE286313.EPICv2.chr1-Y.group_df.qs2"),
  #   "--cores", "1",
  #   "--method", "impute.knn"
  #   , "--debug"
  # )
)
stopifnot(
  argv$method %in% c(
    "knn",
    "knn_tree",
    "knn_tree_chr",
    "pca",
    "missMDA",
    "impute.knn",
    "methyLImp2_fork",
    "methyLImp2_psock"
  )
)
stopifnot(file.exists(argv$beta), file.exists(argv$group))
beta_matrix <- qs2::qs_read(argv$beta)
group_df <- qs2::qs_read(argv$group)
if (argv$debug) {
  message("debug, ", argv$method, " ;cores, ", argv$cores)
  group_df <- subset(group_df, group %in% c("21", "22"))
  beta_matrix <- beta_matrix[, do.call(c, group_df$features)]
  .progress <- TRUE
} else {
  message("full run, ", argv$method, " ;cores, ", argv$cores)
  .progress <- FALSE
}
set.seed(1234)
if (argv$method == "knn") {
  library(slideimp)
  results <- group_imp(
    beta_matrix,
    group = group_df,
    k = 20,
    cores = argv$cores,
    .progress = .progress
  )
} else if (argv$method == "knn_tree") {
  library(slideimp)
  results <- group_imp(
    beta_matrix,
    group = group_df,
    k = 20,
    tree = TRUE,
    cores = argv$cores,
    .progress = .progress
  )
} else if (argv$method == "knn_tree_chr") {
  library(slideimp)
  library(mirai)
  if (argv$cores > 1) {
    daemons(argv$cores)
  }
  results <- group_imp(
    beta_matrix,
    group = group_df,
    k = 20,
    tree = TRUE,
    cores = 1,
    .progress = .progress
  )
  daemons(0)
} else if (argv$method == "pca") {
  library(slideimp)
  library(mirai)
  if (argv$cores > 1) {
    daemons(argv$cores)
  }
  results <- group_imp(
    beta_matrix,
    group = group_df,
    ncp = 10,
    maxiter = 5000,
    .progress = .progress
  )
  daemons(0)
} else if (argv$method == "missMDA") {
  library(missMDA)
  library(furrr)
  options(future.globals.maxSize = 1.0 * 1e9)
  if (argv$cores > 1) {
    plan(multisession, workers = argv$cores)
  }

  results <- future_map(
    group_df$features,
    \(features) {
      imputePCA(
        X = beta_matrix[, features, drop = FALSE],
        ncp = 10,
        maxiter = 5000
      )$completeObs
    },
    .options = furrr_options(seed = TRUE),
    .progress = .progress
  )
  results <- do.call(cbind, results)

  plan(sequential)
} else if (argv$method == "impute.knn") {
  library(impute)
  library(furrr)
  options(future.globals.maxSize = 1.0 * 1e9)
  if (argv$cores > 1) {
    plan(multisession, workers = argv$cores)
  }

  results <- future_map(
    group_df$features,
    \(features) {
      t(
        impute.knn(
          t(beta_matrix[, features, drop = FALSE]),
          k = 20,
          # FULL K-NN.
          maxp = length(features),
          colmax = 0.98,
        )$data
      )
    },
    .options = furrr_options(seed = TRUE),
    .progress = .progress
  )
  results <- do.call(cbind, results)

  plan(sequential)
} else if (argv$method %in% c("methyLImp2_fork", "methyLImp2_psock")) {
  suppressMessages(library(methyLImp2))
  suppressMessages(library(BiocParallel))
  annotation_df <- tidyr::unnest(group_df, cols = features)
  names(annotation_df) <- c("cpg", "chr")
  BPPARAM <- if (argv$cores > 1) {
    if (argv$method == "methyLImp2_psock") {
      SnowParam(workers = argv$cores, type = "SOCK", progressbar = .progress)
    } else {
      MulticoreParam(workers = argv$cores, progressbar = .progress)
    }
  } else {
    SerialParam(progressbar = .progress)
  }
  results <- methyLImp2(
    input = beta_matrix,
    type = "user",
    annotation = annotation_df,
    BPPARAM = BPPARAM
  )
}
```

## Slurm
```{r}
library(tidyverse)

# fork is slower than psock due to R's GC behavior. Just use psock
all_jobs <- tidyr::expand_grid(
  method = c(
    "knn_tree_chr"
    # "knn", "knn_tree_chr", "impute.knn", "missMDA", "pca", "methyLImp2_psock"
  ),
  cores = c(1, 2, 4, 8),
  rep = 1:5
)

all_jobs$output <- with(all_jobs, glue::glue("output/{method}-{cores}-{rep}"))

all_jobs$job <- with(all_jobs, glue::glue(
  '/usr/bin/time -f "%U\\t%S\\t%e\\t%P\\t%M" -o {output}.txt',
  "apptainer exec $APPTAINER/missRanger/missRanger.sif Rscript benchmark.R",
  "--beta input/GSE286313.EPICv2.chr1-Y.qs2 --group input/GSE286313.EPICv2.chr1-Y.group_df.qs2",
  "--cores {cores} --method {method}",
  .sep = " "
))

all_jobs <- tidyr::nest(all_jobs, .by = "cores")

all_jobs$joblist <- here::here(
  "benchmark",
  "multi_cores",
  glue::glue("joblist_{all_jobs$cores}.txt")
)

purrr::walk2(
  all_jobs$data,
  all_jobs$joblist,
  \(x, y) {
    readr::write_lines(x$job, y)
  }
)

max_jobs <- c(30, 10, 5, 5)
max_mem <- c(15, 30, 40, 40)

all_jobs$dsq <- glue::glue(
  "dsq --job-file input/joblist_{all_jobs$cores}.txt --batch-file multi_cores_{all_jobs$cores}.sh",
  "--job-name multi_cores_{all_jobs$cores} --max-jobs {max_jobs}",
  "--output dSQ/dsq-multi_cores_{all_jobs$cores}-%A_%a-%N.out --status-dir .",
  "--time 01:00:00 --nodes 1 --ntasks-per-node 1 --cpus-per-task {all_jobs$cores}",
  "--mem {max_mem}G --partition day",
  .sep = " "
) |>
  as.character()

all_jobs$dsq |> cat(sep = "\n")
```

```{r}
library(here)
library(hhprf)
library(tidyverse)

job_status <- tibble(
  jl = here(
    "benchmark", "multi_cores",
    c(
      "job_6962292_status.tsv"
    )
  ),
  out = here(
    "benchmark", "multi_cores",
    c("joblist_2.txt")
  )
)

map2(job_status$jl, job_status$out, hhprf::retry_dsq)
```

## Viz
Ran the jobs using Slurm and /usr/time. We can use slurm sacc -j to query the actual RSS and Wall Time.

The output/ results are less correct but can be used as backup.

```{r}
library(tidyverse)

multi_cores <- tibble(
  files = list.files("benchmark/multi_cores/dSQ"),
  paths = list.files("benchmark/multi_cores/dSQ", full.names = TRUE)
)

multi_cores$output <- lapply(multi_cores$paths, read_lines)
multi_cores$output |>
  lapply(\(x) {
    grepv("full run", x)
  }) |>
  unique()

extractor <- function(x) {
  method <- if (any(str_detect(x, "knn_tree_chr"))) {
    "knn_tree_chr"
  } else if (any(str_detect(x, "impute.knn"))) {
    "impute.knn"
  } else if (any(str_detect(x, "knn"))) {
    "knn"
  } else if (any(str_detect(x, "pca"))) {
    "pca"
  } else if (any(str_detect(x, "missMDA"))) {
    "missMDA"
  } else if (any(str_detect(x, "methyLImp2_psock"))) {
    "methyLImp2_psock"
  } else if (any(str_detect(x, "methyLImp2_fork"))) {
    "methyLImp2_fork"
  } else {
    # stop("Failed: ", x)
    "remove"
  }
  return(method)
}

multi_cores$method <- vapply(multi_cores$output, extractor, character(1))
multi_cores <- multi_cores |> filter(method != "remove")
multi_cores$cores <- as.numeric(
  str_extract(multi_cores$files, "(?<=multi_cores_)\\d+(?=-)")
)
multi_cores$job_id <- str_extract(multi_cores$files, "(?<=-)[0-9]+_[0-9]+(?=-)")
multi_cores[c("files", "paths", "output")] <- NULL
multi_cores |> print(n = Inf)

cmd <- glue::glue(
  "sacct -j", "{paste0(multi_cores$job_id, collapse = ',')}",
  "--format=JobID,JobName,State,Elapsed,MaxRSS,TotalCPU",
  "--parsable > multi_cores_raw.txt",
  .sep = " "
)

cmd
```

```{r}
library(scales)
multi_cores_raw <- read_delim(
  here::here("benchmark", "multi_cores", "multi_cores_raw.txt"),
  delim = "|",
) |>
  filter(State == "COMPLETED", JobName == "batch") |>
  mutate(
    job_id = str_remove(JobID, ".batch"),
    JobID = NULL,
    State = NULL,
    `...7` = NULL,
    JobName = NULL,
    max_rss_bytes = case_when(
      str_detect(MaxRSS, "K$") ~ readr::parse_number(MaxRSS) * 1024,
      str_detect(MaxRSS, "M$") ~ readr::parse_number(MaxRSS) * 1024^2,
      TRUE ~ readr::parse_number(MaxRSS)
    ),
    MaxRSS = NULL
  )

multi_cores_output <- inner_join(
  multi_cores,
  multi_cores_raw,
  by = "job_id"
)

saveRDS(
  multi_cores_output,
  here::here("benchmark", "multi_cores", "multi_cores_output.rds")
)
```

# Increase Missing
## Tune
```{r tune_imp_mcar.R, eval =  F}
library(argparser)
library(slideimp)
library(qs2)
library(tidyr)
library(fst)
library(mirai)
library(dplyr)
library(glue)

p <- arg_parser("tune_imp")
p <- add_argument(p, "--beta", help = "Path to beta object", type = "character")
p <- add_argument(p, "--output", help = "output file path", type = "character")
p <- add_argument(p, "--tag", help = "tag", type = "character")
p <- add_argument(p, "--cores", help = "n cores", type = "numeric")
p <- add_argument(p, "--global_na", help = "global_na percentage", type = "numeric")
p <- add_argument(p, "--debug", help = "debug mode", flag = TRUE)

argv <- parse_args(
  p
  # ,
  # c(
  #   "--beta", here::here("GSE286313", "EPICv2", "GSE286313.EPICv2_chr22.qs2"),
  #   "--output", here::here("benchmark", "accuracy", "tune"),
  #   "--tag", "v1",
  #   "--cores", "7",
  #   "--global_na", 0.00
  #   # , "--debug"
  # )
)

# input
stopifnot(file.exists(argv$beta))
if (!dir.exists(argv$output)) {
  dir.create(argv$output, recursive = TRUE)
}

beta_name <- tools::file_path_sans_ext(basename(argv$beta))

# knn_imp params
knn_imp_param <- tidyr::expand_grid(
  # v1
  # k = c(10, 15, 20, 25),
  # v2
  k = c(27, 30, 35),
  dist_pow = c(3.5, 4, 4.5),
  method = c("euclidean"),
  post_imp = FALSE,
  cores = argv$cores
)

# pca_imp params. Features in columns
# v1
# pca_imp_param <- tidyr::expand_grid(
#   ncp = c(25, 30, 35),
#   coeff.ridge = c(0.75, 1, 1.25),
#   nb.init = 5,
#   row.w = "n_miss"
# )

# v2
pca_imp_param <- tidyr::expand_grid(
  ncp = c(35, 40),
  coeff.ridge = c(0.9, 1, 1.1),
  nb.init = 5,
  row.w = "n_miss"
)


# input
obj <- qs_read(argv$beta)

# inject global NA level
length_obj <- length(obj)
set.seed(1234)
dim(obj)
obj[sample.int(length_obj, size = ceiling(argv$global_na * length_obj))] <- NA
miss <- is.na(obj)
obj <- obj[(rowSums(miss) / ncol(miss)) < 0.9, ]
message(
  "Actual global missing % is ", round((sum(is.na(obj)) * 100) / length(obj), 2)
)
dim(obj)
if (argv$debug) {
  obj <- obj[1:1000, ]
}
obj_t <- t(obj)
num_na <- if (argv$debug) {
  10
} else {
  5000
}

# tuning
set.seed(1234)
knn_tune_results <- tune_imp(
  obj = obj_t,
  parameters = knn_imp_param,
  .f = "knn_imp",
  rep = 15,
  num_na = num_na,
  cores = argv$cores,
  .progress = TRUE,
  colmax = 0.91
)

# process knn
knn_tune_results <- compute_metrics(knn_tune_results)
write_fst(
  knn_tune_results,
  file.path(
    argv$output,
    glue("knn_imp_tune-{beta_name}-{argv$global_na}-{argv$tag}.fst")
  )
)

# setup parallel
daemons(argv$cores)
set.seed(1234)
pca_imp_tune_results <- tune_imp(
  obj = obj_t,
  parameters = pca_imp_param,
  .f = "pca_imp",
  rep = 15,
  num_na = num_na,
  .progress = TRUE,
  colmax = 0.91
)
# shutdown parallel
daemons(0)
# Process pca_imp
pca_imp_tune_results <- compute_metrics(pca_imp_tune_results)

write_fst(
  pca_imp_tune_results,
  file.path(
    argv$output,
    glue("pca_imp_tune-{beta_name}-{argv$global_na}-{argv$tag}.fst")
  )
)
```

```{bash, eval = F}
#!/bin/bash
#SBATCH --job-name=task
#SBATCH --output=./slurm-%j.out
#SBATCH --time=02:00:00
#SBATCH --nodes=1
#SBATCH --ntasks-per-node=1
#SBATCH --cpus-per-task=30
#SBATCH --mem=30G
#SBATCH --partition=day_amd
#SBATCH --mail-type=END

apptainer exec $APPTAINER/missRanger/missRanger.sif Rscript ./tune_imp_mcar.R \
  --beta "GSE286313/GSE286313.EPICv2_chr22.qs2" \
  --output "benchmark/tune" \
  --tag "v2" \
  --cores 30 \
  --global_na 0
```

```{r}
knn_viz <- read_fst(
  here::here("benchmark", "accuracy", "tune", "knn_imp_tune-GSE286313.EPICv2_chr22-0-v2.fst")
)

knn_viz |>
  ggplot(aes(x = factor(k), color = factor(dist_pow), y = .estimate)) +
  geom_boxplot() +
  facet_wrap(~.metric, scales = "free_y")

pca_viz <- read_fst(
  here::here("benchmark", "accuracy", "tune", "pca_imp_tune-GSE286313.EPICv2_chr22-0-v2.fst")
)

pca_viz |>
  ggplot(aes(x = factor(ncp), color = factor(coeff.ridge), y = .estimate)) +
  geom_boxplot() +
  facet_wrap(~.metric, scales = "free_y")
```

## Fit
Can't inject too much NA because without any cols without any NA, the imputation fails on `{methyLImp2}`

```{r fit_imp_mcar.R}
library(argparser)
library(impute)
library(slideimp)
library(qs2)
library(tidyr)
library(fst)
library(mirai)
library(dplyr)
library(cli)
library(glue)
suppressMessages(library(methyLImp2))

p <- arg_parser("fit_imp_mcar.R")
p <- add_argument(p, "--beta", help = "Path to beta object", type = "character")
p <- add_argument(p, "--output", help = "output file path", type = "character")
p <- add_argument(p, "--cores", help = "n cores", type = "numeric")
p <- add_argument(
  p, "--global_na",
  help = "global_na percentages", type = "numeric", nargs = Inf
)
p <- add_argument(p, "--debug", help = "debug mode", flag = TRUE)
p <- add_argument(p, "--force_knn", help = "overwrite existing knn_imp output", flag = TRUE)
p <- add_argument(p, "--force_pca", help = "overwrite existing pca_imp output", flag = TRUE)
p <- add_argument(p, "--force_methyLImp2", help = "overwrite existing methyLImp2 output", flag = TRUE)

argv <- parse_args(
  p
  # ,
  # c(
  #   # "--beta", here::here("GSE286313", "EPICv2", "GSE286313.EPICv2_chr22.qs2"),
  #   "--beta", here::here("GSE264438", "GSE264438_chr1.qs2"),
  #   "--output", "benchmark",
  #   "--cores", "1",
  #   "--debug",
  #   # "--global_na", c(0, 0.015, 0.02, 0.025, 0.05, 0.10)
  #   "--global_na", c(0, 0.045, 0.05, 0.06, 0.065)
  # )
)

# File and directory validation
stopifnot(file.exists(argv$beta))
if (!dir.exists(argv$output)) {
  dir.create(argv$output, recursive = TRUE)
}

# knn_imp params
knn_imp_param <- tibble::tibble(
  k = 25,
  dist_pow = 4,
  post_imp = FALSE
)

# pca_imp params
pca_imp_param <- tibble::tibble(
  ncp = 40,
  coeff.ridge = 0.9,
  row.w = "n_miss",
  nb.init = 10
)

# methyLImp2 params
methyLImp2_wrap <- function(obj) {
  methyLImp2::methyLImp2(
    obj,
    type = "user",
    annotation = data.frame(cpg = colnames(obj), chr = "custom_chr"),
    BPPARAM = BiocParallel::SerialParam()
  )
}

# input
obj <- t(qs_read(argv$beta))
miss <- is.na(obj)
obj <- obj[, (colSums(miss) / nrow(miss)) < 0.90]

# calculate parameters
total_elements <- length(obj)
current_miss <- sum(is.na(obj))

global_na_percs <- argv$global_na
target_n_miss <- ceiling(total_elements * global_na_percs)
n_miss <- pmax(0, target_n_miss - current_miss)
n_miss <- sort(unique(n_miss))
global_rates <- round((n_miss + current_miss) * 100 / total_elements, 2)
names(n_miss) <- global_rates

seed <- 123456
num_na <- 5000
rep <- 30

# ampute the data
set.seed(seed)
amputed <- lapply(
  n_miss,
  \(i) {
    amp <- obj
    if (i > 0) {
      na_loc <- inject_na(amp, num_na = i, colmax = 0.9)
      amp[na_loc] <- NA
      message("before is: ", current_miss, "; after is: ", sum(is.na(amp)))
    }
    return(amp)
  }
)
names(amputed) <- names(n_miss)

beta_name <- tools::file_path_sans_ext(basename(argv$beta))

# pre-compute output paths
knn_paths <- setNames(
  file.path(argv$output, glue("knn_imp_fitted-{beta_name}-{names(amputed)}.fst")),
  names(amputed)
)
pca_paths <- setNames(
  file.path(argv$output, glue("pca_imp_fitted-{beta_name}-{names(amputed)}.fst")),
  names(amputed)
)
methyLImp2_paths <- setNames(
  file.path(argv$output, glue("methyLImp2_fitted-{beta_name}-{names(amputed)}.fst")),
  names(amputed)
)

daemons(argv$cores)
on.exit(daemons(0), add = TRUE)

for (rate_name in names(amputed)) {
  obj <- amputed[[rate_name]]
  message("Processing global missing rate: ", rate_name, "%")

  # knn_imp
  if (argv$force_knn || !file.exists(knn_paths[[rate_name]])) {
    tryCatch(
      {
        set.seed(seed)
        knn_tune_results <- tune_imp(
          obj = obj,
          parameters = knn_imp_param,
          .f = "knn_imp",
          rep = rep,
          num_na = num_na,
          .progress = FALSE,
          colmax = 0.91
        )
        knn_tune_results <- compute_metrics(knn_tune_results)
        temp_knn <- dplyr::mutate(knn_tune_results, global_na = as.numeric(rate_name))
        write_fst(temp_knn, knn_paths[[rate_name]])
      },
      error = function(e) {
        warning("knn_imp failed for rate ", rate_name, ": ", conditionMessage(e))
      }
    )
  } else {
    message("Skipping knn_imp for rate ", rate_name, "% (output exists)")
  }

  # pca_imp
  if (argv$force_pca || !file.exists(pca_paths[[rate_name]])) {
    tryCatch(
      {
        set.seed(seed)
        pca_imp_tune_results <- tune_imp(
          obj = obj,
          parameters = pca_imp_param,
          .f = "pca_imp",
          rep = rep,
          num_na = num_na,
          .progress = FALSE,
          colmax = 0.91
        )
        pca_imp_tune_results <- compute_metrics(pca_imp_tune_results)
        temp_pca_imp <- dplyr::mutate(pca_imp_tune_results, global_na = as.numeric(rate_name))
        write_fst(temp_pca_imp, pca_paths[[rate_name]])
      },
      error = function(e) {
        warning("pca_imp failed for rate ", rate_name, ": ", conditionMessage(e))
      }
    )
  } else {
    message("Skipping pca_imp for rate ", rate_name, "% (output exists)")
  }

  # methyLImp2
  if (argv$force_methyLImp2 || !file.exists(methyLImp2_paths[[rate_name]])) {
    tryCatch(
      {
        set.seed(seed)
        methyLImp2_tune_results <- tune_imp(
          obj = obj,
          .f = methyLImp2_wrap,
          rep = rep,
          num_na = num_na,
          .progress = FALSE,
          colmax = 0.91
        )
        methyLImp2_tune_results <- compute_metrics(methyLImp2_tune_results)
        temp_methyLImp2 <- dplyr::mutate(methyLImp2_tune_results, global_na = as.numeric(rate_name))
        write_fst(temp_methyLImp2, methyLImp2_paths[[rate_name]])
      },
      error = function(e) {
        warning("methyLImp2 failed for rate ", rate_name, ": ", conditionMessage(e))
      }
    )
  } else {
    message("Skipping methyLImp2 for rate ", rate_name, "% (output exists)")
  }
}
```

```{r}
EPICv2_joblist <- glue::glue(
  "apptainer exec $APPTAINER/missRanger/missRanger.sif Rscript fit_imp_mcar.R",
  "--beta GSE286313/GSE286313.EPICv2_chr22.qs2",
  "--output benchmark",
  "--cores 30",
  "--force_knn",
  "--force_pca",
  "--global_na {c(0, 0.025, 0.05)}",
  .sep = " "
)

readr::write_lines(
  EPICv2_joblist,
  here::here("benchmark", "accuracy", "EPICv2_joblist.txt")
)

MSA_joblist <- glue::glue(
  "apptainer exec $APPTAINER/missRanger/missRanger.sif Rscript fit_imp_mcar.R",
  "--beta GSE264438/GSE264438_chr1.qs2",
  "--output benchmark",
  "--cores 15",
  "--global_na {c(0, 0.045, 0.05, 0.06, 0.065)}",
  .sep = " "
)

readr::write_lines(
  MSA_joblist,
  here::here("benchmark", "accuracy", "MSA_joblist.txt")
)

dsq_gen(
  "EPICv2_joblist.txt",
  batch_file = "EPICv2_fit_imp_mcar.sh",
  job_name = "EPICv2_fit_imp_mcar",
  output = "dSQ/dsq-jobfile-%A_%a-%N.out",
  max_jobs = 3,
  slurm_args = slurm_par(
    time = "04:00:00",
    cpus_per_task = "30",
    mem = "60G"
    # , partition = "day_amd"
  )
)

dsq_gen(
  "MSA_joblist.txt",
  batch_file = "MSA_fit_imp_mcar.sh",
  job_name = "MSA_fit_imp_mcar",
  output = "dSQ/dsq-jobfile-%A_%a-%N.out",
  max_jobs = 2,
  slurm_args = slurm_par(
    time = "12:00:00",
    cpus_per_task = "15",
    mem = "50G"
  )
)

# dsq --job-file EPICv2_joblist.txt --batch-file EPICv2_fit_imp_mcar.sh --job-name EPICv2_fit_imp_mcar --max-jobs 2 --output dSQ/dsq-jobfile-%A_%a-%N.out --status-dir . --time 02:00:00 --nodes 1 --ntasks-per-node 1 --cpus-per-task 15 --mem 50G --partition day

# dsq --job-file MSA_joblist.txt --batch-file MSA_fit_imp_mcar.sh --job-name MSA_fit_imp_mcar --max-jobs 2 --output dSQ/dsq-jobfile-%A_%a-%N.out --status-dir . --time 12:00:00 --nodes 1 --ntasks-per-node 1 --cpus-per-task 15 --mem 50G --partition day
```

Verify if `methyLImp2` works with data where every column has at least one missing value
```{r, eval = FALSE}
library(methyLImp2)
library(qs2)

obj <- t(qs_read(here::here("GSE286313", "EPICv2", "GSE286313.EPICv2_chr22.qs2")))
obj <- obj[, 1:500]
miss <- is.na(obj)
dim(obj)
obj <- obj[, (colSums(miss) / nrow(miss)) < 0.95]
dim(obj)

tryCatch(
  results <- methyLImp2::methyLImp2(
    obj,
    type = "user",
    annotation = data.frame(cpg = colnames(obj), chr = "custom_chr"),
    BPPARAM = BiocParallel::SerialParam()
  )
)

# Inject missing values
complete_cols <- which(colSums(is.na(obj)) == 0)

inject <- cbind(
  row = sample(nrow(obj), length(complete_cols), replace = TRUE),
  col = complete_cols
)

# inject NAs
obj[inject] <- NA

tryCatch(
  results <- methyLImp2::methyLImp2(
    obj,
    type = "user",
    annotation = data.frame(cpg = colnames(obj), chr = "custom_chr"),
    BPPARAM = BiocParallel::SerialParam()
  )
)
```

# `group_imp`
Benchmark accuracy of `group_imp` vs full `knn_imp` and `pca_imp` on the chromosome 1:22 of the MSA chip. Focus on values in just `400` random columns each with `25` values (10,000 missing values to measure RMSE/MAE/RSQ). Do this for multiple reps.

## MSA
```{r}
# Get MSA data
MSA <- get_DNAm("MSA", glue::glue("chr{c(1:22, 'X')}"))
MSA$data <- lapply(MSA$path, qs2::qs_read)
col_order <- colnames(MSA$data[[1]])
MSA$data <- lapply(MSA$data, \(x) x[, col_order])
stopifnot(length(unique(lapply(MSA$data, colnames))) == 1)

MSA <- t(purrr::reduce(MSA$data, rbind))
dim(MSA)
miss <- is.na(MSA)
MSA <- MSA[, (colSums(miss) / nrow(miss)) < 0.9]
dim(MSA)
cv <- col_vars(MSA)
MSA <- MSA[, !(is.na(cv) & dplyr::near(cv, 0))]
dim(MSA)

# Get manifest. Needed for `group_imp`
MSA_manifest <- ilmn_manifest("MSA")
all_probes <- intersect(colnames(MSA), MSA_manifest$feature_id)

# Align
MSA_manifest <- subset(MSA_manifest, feature_id %in% all_probes)
MSA <- MSA[, all_probes]

# Select columns
n_rep <- 30
set.seed(1234)

missing_2d <- tibble(
  rep = seq_len(n_rep),
  IlmnID = lapply(rep, \(x) sample(MSA_manifest$feature_id, size = 400))
) |>
  unnest(cols = "IlmnID") |>
  mutate(
    col_index = match(IlmnID, colnames(MSA)),
    row_index = lapply(IlmnID, \(x) {
      # for each column, 25 values are missing
      sample.int(nrow(MSA), size = 25)
    })
  ) |>
  unnest(cols = "row_index")

na_positions <- missing_2d |>
  nest(.by = "rep") |>
  mutate(location = lapply(
    data, \(x) {
      m <- cbind(x$row_index, x$col_index)
      colnames(m) <- c("row", "col")
      m
    }
  ))

na_positions$subset <- lapply(
  na_positions$data,
  \(x) unique(x$IlmnID)
)

names(MSA_manifest) <- c("features", "group")

na_positions$group_df <- lapply(
  na_positions$subset,
  \(x) group_features(MSA, features_df = MSA_manifest, subset = x)
)

na_positions[, c("data")] <- NULL
if (!dir.exists("benchmark/group_imp/input")) {
  dir.create("benchmark/group_imp/input")
}

qs2::qs_save(na_positions, "benchmark/group_imp/input/MSA_na_positions.qs2")
qs2::qs_save(MSA, "benchmark/group_imp/input/MSA.qs2")
```

## EPICv2
```{r}
# Get EPICv2 data
EPICv2 <- get_DNAm("EPICv2", glue::glue("chr{c(1:22, 'X')}"))
EPICv2$data <- lapply(EPICv2$path, qs2::qs_read)
col_order <- colnames(EPICv2$data[[1]])
EPICv2$data <- lapply(EPICv2$data, \(x) x[, col_order])
stopifnot(length(unique(lapply(EPICv2$data, colnames))) == 1)

EPICv2 <- t(purrr::reduce(EPICv2$data, rbind))
dim(EPICv2)
miss <- is.na(EPICv2)
EPICv2 <- EPICv2[, (colSums(miss) / nrow(miss)) < 0.9]
dim(EPICv2)
cv <- col_vars(EPICv2)
EPICv2 <- EPICv2[, !(is.na(cv) & dplyr::near(cv, 0))]
dim(EPICv2)

# Get manifest. Needed for `group_imp`
EPICv2_manifest <- ilmn_manifest("EPICv2")
all_probes <- intersect(colnames(EPICv2), EPICv2_manifest$feature_id)

# Align
EPICv2_manifest <- subset(EPICv2_manifest, feature_id %in% all_probes)
EPICv2 <- EPICv2[, all_probes]

# Select columns
n_rep <- 30
set.seed(1234)

missing_2d <- tibble(
  rep = seq_len(n_rep),
  IlmnID = lapply(rep, \(x) sample(EPICv2_manifest$feature_id, size = 400))
) |>
  unnest(cols = "IlmnID") |>
  mutate(
    col_index = match(IlmnID, colnames(EPICv2)),
    row_index = lapply(IlmnID, \(x) {
      # for each column, 25 values are missing
      sample.int(nrow(EPICv2), size = 25)
    })
  ) |>
  unnest(cols = "row_index")

na_positions <- missing_2d |>
  nest(.by = "rep") |>
  mutate(location = lapply(
    data, \(x) {
      m <- cbind(x$row_index, x$col_index)
      colnames(m) <- c("row", "col")
      m
    }
  ))

na_positions$subset <- lapply(
  na_positions$data,
  \(x) unique(x$IlmnID)
)

names(MSA_manifest) <- c("features", "group")

na_positions$group_df <- lapply(
  na_positions$subset,
  \(x) group_features(EPICv2, features_df = EPICv2_manifest, subset = x)
)

na_positions[, c("data")] <- NULL
if (!dir.exists("benchmark/group_imp/input")) {
  dir.create("benchmark/group_imp/input")
}

qs2::qs_save(na_positions, "benchmark/group_imp/input/EPICv2_na_positions.qs2")
qs2::qs_save(EPICv2, "benchmark/group_imp/input/EPICv2.qs2")
```

## `group_imp.R`
```{r, group_imp.R}
library(argparser)
library(tidyverse)
library(slideimp)
library(fst)
library(glue)
library(qs2)

# Set up argument parser
p <- arg_parser("Run group/full imputation with knn/pca method")
p <- add_argument(p, "--na_positions", help = "Path to na_positions object")
p <- add_argument(p, "--beta", help = "Path to beta object")
p <- add_argument(p, "--output", help = "Output path")
p <- add_argument(p, "--rep", help = "Repeat number", type = "integer")
p <- add_argument(p, "--method", help = "Method: pca or knn")
p <- add_argument(p, "--group", help = "Group type: group or full")

argv <- parse_args(
  p,
  c(
    "--na_positions", "benchmark/group_imp/input/MSA_na_positions.qs2",
    "--beta", "benchmark/group_imp/input/MSA.qs2",
    "--output", "benchmark/group_imp/output",
    "--rep", 1,
    "--method", "knn",
    "--group", "group"
  )
)

# Check that paths exist
stopifnot(file.exists(argv$na_position))
stopifnot(file.exists(argv$beta))
stopifnot(dir.exists(argv$output))

# Load data
na_positions <- qs_read(argv$na_position)
beta <- qs_read(argv$beta)

# Output file path
results <- file.path(
  argv$output,
  glue::glue("{argv$method}-{argv$group}-{argv$rep}.fst")
)

# Set parameters
k <- 25
ncp <- 25

set.seed(12345)

# Wrapper functions
group_imp_knn_wrap <- function(obj, group_df) {
  imputed <- group_imp(
    obj = obj,
    k = k,
    group = group_df
  )[, colnames(obj)]
  return(imputed)
}

group_imp_pca_wrap <- function(obj, group_df) {
  imputed <- group_imp(
    obj = obj,
    ncp = ncp,
    group = group_df
  )[, colnames(obj)]
  return(imputed)
}

knn_imp_wrap <- function(obj, subset) {
  imputed <- knn_imp(
    obj = obj,
    k = k,
    subset = subset
  )
  obj[, subset] <- imputed[, subset]
  return(obj)
}

pca_imp_wrap <- function(obj, subset) {
  imputed <- pca_imp(
    obj = obj,
    ncp = ncp
  )[, colnames(obj)]
  return(imputed)
}

# Get the data for this specific rep
i <- argv$rep
message("Processing rep ", i)

# Get the position of missing for `tune_imp`
position <- list(na_positions$location[[i]])
# Get the column names for this rep
subset <- list(na_positions$subset[[i]])
# Get the group_df for this rep
group_df <- list(na_positions$group_df[[i]])

# Run the appropriate combination based on method and group
if (argv$method == "knn" && argv$group == "full") {
  message("Running KNN Full")
  tune_result <- tune_imp(
    obj = beta,
    parameters = tibble::tibble(subset = subset),
    .f = knn_imp_wrap,
    rep = position
  )
  imputed <- tune_result$result[[1]]
} else if (argv$method == "knn" && argv$group == "group") {
  message("Running KNN Group")
  tune_result <- tune_imp(
    obj = beta,
    parameters = tibble::tibble(group_df = group_df),
    .f = group_imp_knn_wrap,
    rep = position
  )
  imputed <- tune_result$result[[1]]
} else if (argv$method == "pca" && argv$group == "full") {
  message("Running PCA Full")
  tune_result <- tune_imp(
    obj = beta,
    parameters = tibble::tibble(subset = subset),
    .f = pca_imp_wrap,
    rep = position
  )
  imputed <- tune_result$result[[1]]
} else if (argv$method == "pca" && argv$group == "group") {
  message("Running PCA Group")
  tune_result <- tune_imp(
    obj = beta,
    parameters = tibble::tibble(group_df = group_df),
    .f = group_imp_pca_wrap,
    rep = position
  )
  imputed <- tune_result$result[[1]]
} else {
  stop("Method must be 'knn' or 'pca', group must be 'full' or 'group'")
}

warnings()

# Save results
message("Saving results to ", results)
write_fst(imputed, results)
message("Done!")
```

## dSQ
```{r}
platform <- "MSA"
# platform <- "EPICv2"

joblist_tbl <- tidyr::expand_grid(
  "na_positions" = glue::glue("input/{platform}_na_positions.qs2"),
  "beta" = glue::glue("input/{platform}.qs2"),
  "output" = glue::glue("output/{platform}"),
  "rep" = 1:30,
  "method" = c("pca", "knn"),
  "group" = c("group", "full")
)

joblist_tbl$joblist <- with(
  joblist_tbl,
  glue::glue(
    "apptainer exec $APPTAINER/missRanger/missRanger.sif Rscript group_imp.R",
    "--na_positions", "{na_positions}",
    "--beta", "{beta}",
    "--output", "{output}",
    "--rep", "{rep}",
    "--method", "{method}",
    "--group", "{group}",
    .sep = " "
  )
)

readr::write_lines(
  joblist_tbl$joblist,
  glue::glue("benchmark/group_imp/{platform}_joblist.txt")
)

dsq_gen(
  glue::glue("{platform}_joblist.txt"),
  batch_file = "group_imp.sh",
  job_name = "group_imp",
  output = "dSQ/dsq-jobfile-%A_%a-%N.out",
  max_jobs = 40,
  slurm_args = slurm_par(
    time = "06:00:00",
    cpus_per_task = "1",
    mem = "20G"
  )
)

# dsq --job-file joblist.txt --batch-file group_imp.sh --job-name group_imp --max-jobs 30 --output dSQ/dsq-jobfile-%A_%a-%N.out --status-dir . --time 12:00:00 --nodes 1 --ntasks-per-node 1 --cpus-per-task 1 --mem 30G --partition day
```

## Viz
```{r}
# platform <- "MSA"
platform <- "EPICv2"

group_imp <- tibble(
  files = list.files(
    glue::glue("benchmark/group_imp/output/{platform}")
  ),
  paths = list.files(
    glue::glue("benchmark/group_imp/output/{platform}"),
    full.names = TRUE
  )
) |>
  separate_wider_delim(
    cols = "files",
    delim = "-", names = c("method", "group", "rep")
  ) |>
  mutate(result = lapply(paths, fst::read_fst))

class(group_imp) <- c("TuneImp", "tbl_df", "tbl", "data.frame")

group_imp <- compute_metrics(group_imp, c("mae", "rmse", "rsq", "bias"))
group_imp$paths <- NULL
group_imp$rep <- as.numeric(str_remove(group_imp$rep, ".fst"))

fst::write_fst(group_imp, glue::glue("benchmark/{platform}_full_vs_group.fst"))
```

# impute.knn and ball tree
* impute.knn maxp and ball tree have quirks to be showcase here
```{r, impute.knn.R}
library(argparser)
library(impute)
library(slideimp)
library(qs2)
library(tibble)
library(tidyr)
library(dplyr)
library(cli)
library(mirai)

p <- arg_parser("impute.knn and ball tree")
p <- add_argument(p, "--beta", help = "Path to beta object")
p <- add_argument(p, "--output", help = "Output path")
p <- add_argument(p, "--cores", help = "Cores", type = "numeric")

argv <- parse_args(
  p
  # ,
  # c(
  #   "--beta", here::here("GSE286313", "EPICv2", "GSE286313.EPICv2_chr22.qs2"),
  #   "--output", here::here("benchmark", "impute.knn"),
  #   "--cores", 2
  # )
)

stopifnot(file.exists(argv$beta))
if (!dir.exists(argv$output)) {
  dir.create(argv$output)
}

# read in beta
obj <- t(qs_read(argv$beta))
miss <- is.na(obj)
# filter
obj <- obj[, (colSums(miss) / nrow(miss)) < 0.90]

# calculate parameters
total_elements <- length(obj)
current_miss <- sum(is.na(obj))

# global missing rates
rates <- seq(0, 0.2, by = 0.05)
target_n_miss <- ceiling(total_elements * rates)
# how many additional missing values we need to inject
n_miss <- pmax(0, target_n_miss - current_miss)
n_miss <- sort(unique(n_miss))
global_rates <- round((n_miss + current_miss) * 100 / total_elements, 1)
names(n_miss) <- global_rates

# ampute the data
sample_from <- which(!is.na(obj))

set.seed(1234)
amputed <- lapply(
  n_miss,
  \(i) {
    amp <- obj
    if (i > 0) {
      na_loc <- inject_na(amp, num_na = i, colmax = 0.9)
      amp[na_loc] <- NA
      message("before is: ", current_miss, "; after is: ", sum(is.na(amp)))
    }
    return(amp)
  }
)
names(amputed) <- names(n_miss)

# maxp of impute.knn converts to how many columns
maxp <- ceiling(
  pmax(
    1500,
    seq(0, 1, by = 0.1) * ncol(obj)
  )
)
maxp_prop <- seq(0, 1, by = 0.1)

# testing parameters
parameters <- tibble(rate = names(n_miss[1]), maxp = maxp, method = "impute.knn") |>
  add_row(
    expand_grid(
      rate = names(n_miss),
      maxp = c(ceiling(0.5 * ncol(obj)), ncol(obj)),
      method = c("impute.knn", "knn_imp", "knn_imp_tree")
    )
  ) |>
  mutate(maxp = if_else(method != "impute.knn", NA_real_, maxp)) |>
  # unique because the expand_grid created an extra same row for impute.knn
  unique() |>
  left_join(tibble(maxp = maxp, maxp_prop = maxp_prop), by = "maxp")

parameters <- parameters[sample.int(nrow(parameters)), ]

impute_knn_wrapper <- function(obj, maxp) {
  suppressMessages(t(impute::impute.knn(data = t(obj), k = 20, colmax = 0.91, maxp = maxp)$data))
}

results <- list()

daemons(argv$cores)
tryCatch({
  cli_progress_bar(total = nrow(parameters))
  for (i in seq_len(nrow(parameters))) {
    cli_progress_update(set = i, status = parameters$method[i])
    if (parameters$method[i] == "impute.knn") {
      tune <- tune_imp(
        obj = amputed[[parameters$rate[i]]],
        parameters = data.frame(maxp = parameters$maxp[i]),
        .f = impute_knn_wrapper,
        rep = 30,
        num_na = 5000,
        colmax = 0.9,
        .progress = FALSE
      )
    } else if (parameters$method[i] == "knn_imp") {
      tune <- tune_imp(
        obj = amputed[[parameters$rate[i]]],
        parameters = data.frame(k = 20),
        .f = "knn_imp",
        rep = 30,
        num_na = 5000,
        colmax = 0.9,
        .progress = FALSE
      )
    } else if (parameters$method[i] == "knn_imp_tree") {
      tune <- tune_imp(
        obj = amputed[[parameters$rate[i]]],
        parameters = data.frame(k = 20, tree = TRUE),
        .f = "knn_imp",
        rep = 30,
        num_na = 5000,
        colmax = 0.9,
        .progress = FALSE
      )
    } else {
      stop("Error")
    }
    tune <- compute_metrics(tune)
    results[[i]] <- dplyr::select(tune, rep, .metric, .estimate)
  }
  cli_progress_done()
}, finally = {
  daemons(0)
})

parameters$results <- results
parameters <- unnest(parameters, cols = "results")
parameters$rate <- as.numeric(parameters$rate)
beta_name <- tools::file_path_sans_ext(basename(argv$beta))

readr::write_csv(parameters, file.path(argv$output, glue::glue("{beta_name}_results.csv")))
```

```{bash}
#!/bin/bash
#SBATCH --job-name=task
#SBATCH --output=./slurm-%j.out
#SBATCH --time=02:00:00
#SBATCH --nodes=1
#SBATCH --ntasks-per-node=1
#SBATCH --cpus-per-task=10
#SBATCH --mem=30G
#SBATCH --partition=day
#SBATCH --mail-type=END

# Execute the bash script within the container
apptainer exec $APPTAINER/missRanger/missRanger.sif Rscript ./impute.knn.R \
  --beta input/GSE286313.EPICv2_chr22.qs2 \
  --output output \
  --cores 10
```

