---
title: "Benchmark"
author: Hung
date: "`r Sys.Date()`"
format:
  html:
    fig-align: center
    toc: true
    toc-depth: 2
    fig-height: 4
    fig-width: 6
    number-sections: true
    toc-location: left
    page-layout: full
execute:  
  echo: false  
  warning: false
  message: false
---

# Synopsis
This file performs the benchmarks

# Scenario
```{r}
#| eval: true
# Install packages if needed
pkgs <- list(
  cran = c("missMDA", "softImpute", "missRanger"),
  bioc = c("methyLImp2", "impute"),
  github = c("hhp94/slideimp", "hhp94/slideimp.extra")
)

if (!rlang::is_installed("BiocManager")) install.packages("BiocManager")

for (pkg in pkgs$cran) {
  if (!rlang::is_installed(pkg)) install.packages(pkg)
}

for (pkg in pkgs$bioc) {
  if (!rlang::is_installed(pkg)) BiocManager::install(pkg)
}

for (pkg in pkgs$github) {
  pkg_name <- sub(".*/", "", pkg)
  if (!rlang::is_installed(pkg_name)) remotes::install_github(pkg)
}

# Load libraries
library(impute)
library(slideimp)
library(bench)
library(here)
library(qs2)
library(duckdb)
library(ggplot2)
library(tidyverse)
library(slideimp.extra)
source("utils.R")
```

# Cleaning
```{r, eval = F}
EPIC_chr1 <- do.call(rbind, lapply(get_DNAm("EPICv1", c("chr1"))$path, qs_read))
EPIC_chr1_miss <- is.na(EPIC_chr1)
EPIC_chr1 <- EPIC_chr1[(rowSums(EPIC_chr1_miss) / ncol(EPIC_chr1_miss)) < 0.90, ]
EPIC_chr1 <- EPIC_chr1[
  which(!rownames(EPIC_chr1) %in% c("cg00391452", "cg02434099")),
]
stopifnot(all(!c("cg00391452", "cg02434099") %in% rownames(EPIC_chr1)))

# Transposed version
EPIC_chr1_t <- t(EPIC_chr1)
```

```{r, eval = F}
f_impute.knn <- function(data, n_samples, k = 20, prop = 0.55, seed = 42) {
  set.seed(seed)
  if (n_samples > nrow(data)) {
    n_samples <- nrow(data)
  }

  selected_rows <- sample.int(nrow(data), size = n_samples)
  maxp <- if (nrow(data) <= 1500) {
    1500
  } else {
    ceiling(prop * nrow(data))
  }
  suppressWarnings(impute.knn(data[selected_rows, ], k = k, maxp = maxp))
}

knn_helper <- function(data_t, n_samples, seed = 42) {
  set.seed(seed)
  if (n_samples > ncol(data_t)) n_samples <- ncol(data_t)
  selected_cols <- sample.int(ncol(data_t), size = n_samples)
  data_t[, selected_cols]
}

f_knn_brute <- function(
  data_t,
  n_samples,
  k = 20,
  seed = 42,
  weighted = FALSE,
  dist_pow = 1,
  cores = 1
) {
  sampled_data <- knn_helper(data_t, n_samples, seed)

  knn_imp(
    sampled_data,
    k = k,
    cores = cores,
    post_imp = TRUE,
    dist_pow = dist_pow
  )
}

f_knn_ball <- function(
  data_t,
  n_samples,
  k = 20,
  seed = 42,
  weighted = FALSE,
  cores = 1,
  dist_pow = 1
) {
  sampled_data <- knn_helper(data_t, n_samples, seed)

  knn_imp(
    sampled_data,
    k = k,
    tree = "ball",
    post_imp = TRUE,
    cores = cores,
    dist_pow = dist_pow
  )
}

f_pca <- function(
  data_t,
  n_samples,
  seed = 42
) {
  sampled_data <- knn_helper(data_t, n_samples, seed)

  pca_imp(sampled_data, ncp = 10, seed = seed)
}

library(methyLImp2)

f_methyLImp2 <- function(data_t, n_samples, seed = 42) {
  sampled_data <- knn_helper(data_t, n_samples, seed)
  methyLImp2(
    input = sampled_data,
    type = "user",
    annotation = data.frame(cpg = colnames(sampled_data), chr = "chr1"),
    BPPARAM = BiocParallel::SerialParam()
  )
}

library(missMDA)

f_imputePCA <- function(data_t, n_samples, seed = 42) {
  sampled_data <- knn_helper(data_t, n_samples, seed)

  imputePCA(X = sampled_data, ncp = 10)$completeObs
}
```

# Speed
```{r, eval = F}
EPIC_500 <- mark(
  f_impute.knn(EPIC_chr1, 500),
  f_methyLImp2(EPIC_chr1_t, 500),
  f_imputePCA(EPIC_chr1_t, 500),
  f_pca(EPIC_chr1_t, 500),
  f_knn_brute(EPIC_chr1_t, 500),
  f_knn_brute(EPIC_chr1_t, 500, cores = 4),
  f_knn_ball(EPIC_chr1_t, 500, cores = 4),
  min_iterations = 5,
  memory = FALSE,
  check = F,
  filter_gc = FALSE
)

EPIC_1000 <- mark(
  f_impute.knn(EPIC_chr1, 1000),
  f_methyLImp2(EPIC_chr1_t, 1000),
  f_imputePCA(EPIC_chr1_t, 1000),
  f_pca(EPIC_chr1_t, 1000),
  f_knn_brute(EPIC_chr1_t, 1000),
  f_knn_brute(EPIC_chr1_t, 1000, cores = 4),
  f_knn_ball(EPIC_chr1_t, 1000, cores = 4),
  min_iterations = 5,
  memory = FALSE,
  check = F,
  filter_gc = FALSE
)

EPIC_5000 <- mark(
  f_impute.knn(EPIC_chr1, 5000),
  f_methyLImp2(EPIC_chr1_t, 5000),
  f_imputePCA(EPIC_chr1_t, 5000),
  f_pca(EPIC_chr1_t, 5000),
  f_knn_brute(EPIC_chr1_t, 5000),
  f_knn_brute(EPIC_chr1_t, 5000, cores = 4),
  f_knn_ball(EPIC_chr1_t, 5000, cores = 4),
  min_iterations = 5,
  memory = FALSE,
  check = F,
  filter_gc = FALSE
)

EPIC_10000 <- mark(
  f_impute.knn(EPIC_chr1, 10000),
  f_methyLImp2(EPIC_chr1_t, 10000),
  f_imputePCA(EPIC_chr1_t, 10000),
  f_pca(EPIC_chr1_t, 10000),
  f_knn_brute(EPIC_chr1_t, 10000),
  f_knn_brute(EPIC_chr1_t, 10000, cores = 4),
  f_knn_ball(EPIC_chr1_t, 10000, cores = 4),
  min_iterations = 5,
  memory = FALSE,
  check = F,
  filter_gc = FALSE
)

# Skip memory for below tests due to size
EPIC_20000 <- mark(
  f_impute.knn(EPIC_chr1, 20000),
  f_methyLImp2(EPIC_chr1_t, 20000),
  f_imputePCA(EPIC_chr1_t, 20000),
  f_pca(EPIC_chr1_t, 20000),
  f_knn_brute(EPIC_chr1_t, 20000),
  f_knn_brute(EPIC_chr1_t, 20000, cores = 4),
  f_knn_ball(EPIC_chr1_t, 20000, cores = 4),
  min_iterations = 5,
  memory = FALSE,
  check = F,
  filter_gc = FALSE
)

EPIC_50000 <- mark(
  f_impute.knn(EPIC_chr1, 50000),
  f_methyLImp2(EPIC_chr1_t, 50000),
  f_imputePCA(EPIC_chr1_t, 50000),
  f_pca(EPIC_chr1_t, 50000),
  f_knn_brute(EPIC_chr1_t, 50000),
  f_knn_brute(EPIC_chr1_t, 50000, cores = 4),
  f_knn_ball(EPIC_chr1_t, 50000, cores = 4),
  min_iterations = 5,
  memory = FALSE,
  check = F,
  filter_gc = FALSE
)

EPIC_72000 <- mark(
  f_impute.knn(EPIC_chr1, 72000),
  f_methyLImp2(EPIC_chr1_t, 72000),
  f_imputePCA(EPIC_chr1_t, 72000),
  f_pca(EPIC_chr1_t, 72000),
  f_knn_brute(EPIC_chr1_t, 72000),
  f_knn_brute(EPIC_chr1_t, 72000, cores = 4),
  f_knn_ball(EPIC_chr1_t, 72000, cores = 4),
  min_iterations = 5,
  memory = FALSE,
  check = F,
  filter_gc = FALSE
)
```

## Export
```{r, eval = F}
walk2(
  list(
    EPIC_500,
    EPIC_1000,
    EPIC_5000,
    EPIC_10000,
    EPIC_20000,
    EPIC_50000,
    EPIC_72000
  ),
  glue::glue("benchmark/EPIC_{c(500, 1000, 5000, 10000, 20000, 50000, 72000)}.qs2"),
  \(x, y) {
    x$memory <- NULL
    qs2::qs_save(x, y)
  }
)
```

```{r, eval = F}
dir_path <- "benchmark/gnu_time"

if (!dir.exists(dir_path)) {
  dir.create(dir_path, recursive = TRUE)
}

saveRDS(
  list(
    f_impute.knn = f_impute.knn,
    knn_helper = knn_helper,
    f_knn_brute = f_knn_brute,
    f_knn_ball = f_knn_ball,
    f_pca = f_pca,
    f_methyLImp2 = f_methyLImp2,
    f_imputePCA = f_imputePCA
  ),
  file.path(dir_path, "functions.rds")
)

saveRDS(
  EPIC_chr1_t,
  file.path(dir_path, "EPIC_chr1_t.rds")
)
```

# Memory
The `bench::mark` tool may not be measuring memory profile of underlying R codes accurately. So we profile peak memory usage using gnu time. The HPC has different nodes but peak memory useage should be the same across all of them.

## Script
```{r memory.R, eval =  F}
library(argparser)

p <- arg_parser("memory profile")
p <- add_argument(
  p,
  "--rds",
  help = "rds file containing the functions",
  type = "character"
)
p <- add_argument(p, "--chr1", help = "chr1 data", type = "character")
p <- add_argument(p, "--cpgs", help = "n CpGs", type = "numeric")
p <- add_argument(p, "--fun_name", help = "fun_name", type = "character")
p <- add_argument(p, "--cores", help = "n cores", type = "numeric")
p <- add_argument(p, "--baseline", help = "baseline", flag = TRUE)

argv <- parse_args(
  p
  # ,
  # c(
  #   "--rds", "benchmark/gnu_time/functions.rds",
  #   "--chr1", "benchmark/gnu_time/EPIC_chr1_t.rds",
  #   "--fun_name", "f_pca",
  #   "--cpgs", 500,
  #   "--cores", 4
  # )
)

stopifnot(file.exists(argv$rds))
stopifnot(file.exists(argv$chr1))

functions <- readRDS(argv$rds)
EPIC_chr1_t <- readRDS(argv$chr1)
EPIC_chr1 <- t(EPIC_chr1_t)

for (i in seq_along(functions)) {
  assign(names(functions)[i], functions[[i]])
}

stopifnot(argv$fun_name %in% names(functions))

set.seed(42)

message(argv$fun_name)

if (!argv$baseline) {
  if (argv$fun_name == "f_impute.knn") {
    library(impute)
    results <- do.call(argv$fun_name, list(data = EPIC_chr1, n_samples = argv$cpgs))
  } else if (argv$fun_name %in% c("f_knn_brute", "f_knn_ball")) {
    library(slideimp)
    results <- do.call(
      argv$fun_name,
      list(
        data = EPIC_chr1_t,
        n_samples = argv$cpgs,
        cores = argv$cores
      )
    )
  } else if (argv$fun_nam == "f_methyLImp2") {
    library(BiocParallel)
    library(methyLImp2)
    results <- do.call(
      argv$fun_name,
      list(data = EPIC_chr1_t, n_samples = argv$cpgs)
    )
  } else if (argv$fun_nam == "f_imputePCA") {
    library(missMDA)
    results <- do.call(
      argv$fun_name,
      list(data = EPIC_chr1_t, n_samples = argv$cpgs)
    )
  } else {
    library(slideimp)
    results <- do.call(
      argv$fun_name,
      list(data = EPIC_chr1_t, n_samples = argv$cpgs)
    )
  }
}
```

## Slurm
```{r}
library(tidyverse)

knn_params <- tidyr::expand_grid(
  f = c("f_knn_brute", "f_knn_ball"),
  cpgs = c(10000, 20000, 50000, 72000),
  cores = c(1, 4),
  rep = 1:5
) |>
  filter(!(cores == 1 & f == "f_knn_ball"))

other_params <- tidyr::expand_grid(
  f = c("f_impute.knn", "f_pca", "f_methyLImp2", "f_imputePCA"),
  cpgs = c(10000, 20000, 50000, 72000),
  cores = 1,
  rep = 1:5
)

all_jobs <- bind_rows(knn_params, other_params)
all_jobs$output <- with(all_jobs, glue::glue("output/{f}-{cpgs}-{cores}-{rep}"))
all_jobs$job <- with(all_jobs, glue::glue(
  '/usr/bin/time -f "%U\\t%S\\t%e\\t%P\\t%M" -o {output}.txt',
  "apptainer exec $APPTAINER/missRanger/missRanger.sif Rscript memory.R",
  "--rds input/functions.rds --chr1 input/EPIC_chr1_t.rds",
  "--fun_name {f} --cpgs {cpgs} --cores {cores}",
  .sep = " "
))

# /usr/bin/time -f "%U\t%S\t%e\t%P\t%M" -o {output} {command}
# %U - User CPU time (seconds)
# %S - System CPU time (seconds)
# %e - Elapsed real time (seconds)
# %P - CPU percentage
# %M - Maximum resident set size (KB)

splitIndices1 <- function(nx, ncl) {
  nx <- as.integer(nx)
  ncl <- as.integer(ncl)
  stopifnot(`nx and ncl must be non negative` = nx > 0 & ncl > 0)
  div <- nx %/% ncl
  mod <- nx %% ncl
  i <- seq_len(nx)
  r <- sort(c(rep(seq_len(ncl), each = div), seq_len(mod)))
  return(unname(split(i, f = factor(r))))
}

jobs <- splitIndices1(length(all_jobs$job), 28)
jobs_list <- lapply(jobs, \(x) paste(all_jobs$job[x], collapse = "; ")) |>
  purrr::list_c()

readr::write_lines(jobs_list, "benchmark/gnu_time/joblist.txt")

dsq_gen(
  "joblist.txt",
  batch_file = "memory_profile.sh",
  job_name = "memory_profile",
  output = "dSQ/dsq-jobfile-%A_%a-%N.out",
  max_jobs = 14,
  slurm_args = slurm_par(
    time = "01:00:00",
    cpus_per_task = "4",
    mem = "40G"
  )
)

# dsq --job-file joblist.txt --batch-file memory_profile.sh --job-name memory_profile --max-jobs 14 --output dSQ/dsq-jobfile-%A_%a-%N.out --status-dir . --time 01:00:00 --nodes 1 --ntasks-per-node 1 --cpus-per-task 4 --mem 40G --partition day
```

## Viz
```{r}
library(tidyverse)

if (!file.exists("benchmark/gnu_time/memory.csv")) {
  memory <- tibble(
    files = list.files("benchmark/gnu_time/output"),
    paths = list.files("benchmark/gnu_time/output", full.names = TRUE)
  ) |>
    separate_wider_delim(files, names = c("type", "cpgs", "cores", "rep"), delim = "-") |>
    filter(type != "baseline") |>
    mutate(
      cpgs = as.numeric(cpgs),
      cores = as.numeric(cores),
      rep = as.numeric(str_remove(rep, ".txt")),
      files = lapply(paths, \(x) suppressMessages(read_tsv(x, col_names = FALSE))),
      paths = NULL
    ) |>
    unnest(cols = files)

  names(memory)[match(paste0("X", 1:5), names(memory))] <- c(
    "user_cpu",
    "system_cpu",
    "elapsed_real_time",
    "cpu_useage",
    "max_resident_set_size_kb"
  )

  memory$cpu_useage <- as.numeric(str_remove(memory$cpu_useage, "%"))
  write_csv(memory, "benchmark/gnu_time/memory.csv")
} else {
  memory <- read_csv("benchmark/gnu_time/memory.csv")
}

viz_df <- filter(memory, type != "baseline") |>
  pivot_longer(
    cols = c(user_cpu, system_cpu, elapsed_real_time, cpu_useage, max_resident_set_size_kb)
  ) |>
  mutate(type = if_else(cores != 1, glue::glue("{type}_{cores}"), type)) |>
  summarize(median_value = median(value), .by = c("type", "cpgs", "cores", "name"))

viz_df |>
  filter(name == "max_resident_set_size_kb", type != "f_knn_brute") |>
  mutate(median_value = median_value * 1024) |>
  ggplot(aes(x = cpgs, y = median_value)) +
  geom_point(alpha = 0.75, aes(color = type), size = 2) +
  geom_line(aes(group = type, color = type)) +
  scale_y_continuous(labels = scales::label_bytes())
```

# Multi Cores
* R2 requests that the benchmark done on multiple cores. We perform grouped imputation on the EPICv2 GSE286313 data using 4/8/12 cores in parallel.
* Scenarios will be K-NN, pca_imp, methyLImp2 PSOCK, and methyLImp2 FORK.
* We will also have to pin the jobs to a specific node for maximum reproducibility.

## Prep
```{r}
library(slideimp.extra)

beta_matrix <- do.call(
  rbind,
  lapply(
    get_DNAm(
      "EPICv2",
      # c(glue::glue("chr{1:8}"))
      c(glue::glue("all"))
    )$path,
    qs_read
  )
) |>
  t()

# beta_matrix <- dedup_matrix(beta_matrix, "EPICv2")
cv <- col_vars(beta_matrix)

beta_matrix |> dim()
beta_matrix <- beta_matrix[, !is.na(cv) & cv > 0]
beta_matrix |> dim()
beta_matrix <- beta_matrix[, (colSums(is.na(beta_matrix)) / nrow(beta_matrix)) < 0.9]
beta_matrix |> dim()

group_df <- group_features(beta_matrix, ilmn_manifest("EPICv2", dedupped = FALSE))
group_df <- dplyr::filter(group_df, !group %in% c("0", "M"))

qs2::qs_save(beta_matrix, here("GSE286313", "GSE286313.EPICv2.chr1-Y.qs2"))
qs2::qs_save(group_df, here("GSE286313", "GSE286313.EPICv2.chr1-Y.group_df.qs2"))
```

## Script
```{r benchmark.R, eval = F}
library(argparser)

p <- arg_parser("benchmark")
p <- add_argument(p, "--beta", help = "beta matrix file path", type = "character")
p <- add_argument(p, "--group", help = "group file path", type = "character")
p <- add_argument(p, "--cores", help = "number of cores", type = "integer")
p <- add_argument(
  p, "--method",
  help = "knn, knn_tree, pca, missMDA, impute.knn, methyLImp2_fork, methyLImp2_psock",
  type = "character"
)
p <- add_argument(p, "--debug", help = "enable debug mode", flag = TRUE)
argv <- parse_args(
  p
  ,
  c(
  "--beta", here::here("GSE286313", "GSE286313.EPICv2.chr1-Y.qs2"),
  "--group", here::here("GSE286313", "GSE286313.EPICv2.chr1-Y.group_df.qs2"),
  "--cores", "1",
  "--method", "pca"
  # , "--debug"
  )
)
stopifnot(
  argv$method %in% c("knn", "knn_tree", "pca", "missMDA", "impute.knn", "methyLImp2_fork", "methyLImp2_psock")
)
stopifnot(file.exists(argv$beta), file.exists(argv$group))
beta_matrix <- qs2::qs_read(argv$beta)
group_df <- qs2::qs_read(argv$group)
if (argv$debug) {
  message("debug, ", argv$method, " ;cores, ", argv$cores)
  group_df <- subset(group_df, group %in% c("21", "22"))
  beta_matrix <- beta_matrix[, do.call(c, group_df$features)]
  .progress <- TRUE
} else {
  message("full run, ", argv$method, " ;cores, ", argv$cores)
  .progress <- FALSE
}
set.seed(1234)
if (argv$method == "knn") {
  library(slideimp)
  results <- group_imp(
    beta_matrix,
    group = group_df,
    k = 20,
    cores = argv$cores,
    .progress = .progress
  )
} else if (argv$method == "knn_tree") {
  library(slideimp)
  results <- group_imp(
    beta_matrix,
    group = group_df,
    k = 20,
    tree = "ball",
    cores = argv$cores,
    .progress = .progress
  )
} else if (argv$method == "pca") {
  library(slideimp)
  library(mirai)
  if (argv$cores > 1) {
    daemons(argv$cores)
  }
  results <- group_imp(
    beta_matrix,
    group = group_df,
    ncp = 10,
    maxiter = 5000,
    .progress = .progress
  )
  daemons(0)
} else if (argv$method == "missMDA") {
  library(missMDA)
  library(furrr)
  options(future.globals.maxSize = 1.0 * 1e9)
  if (argv$cores > 1) {
    plan(multisession, workers = argv$cores)
  }

  results <- future_map(
    group_df$features,
    \(features) imputePCA(
      X = beta_matrix[, features, drop = FALSE],
      ncp = 10,
      maxiter = 5000
    )$completeObs,
    .options = furrr_options(seed = TRUE),
    .progress = .progress
  )
  results <- do.call(cbind, results)

  plan(sequential)
} else if (argv$method == "impute.knn") {
  library(impute)
  library(furrr)
  options(future.globals.maxSize = 1.0 * 1e9)
  if (argv$cores > 1) {
    plan(multisession, workers = argv$cores)
  }

  results <- future_map(
    group_df$features,
    \(features) t(
      impute.knn(
        t(beta_matrix[, features, drop = FALSE]),
        k = 20,
        colmax = 0.98,
      )$data
    ),
    .options = furrr_options(seed = TRUE),
    .progress = .progress
  )
  results <- do.call(cbind, results)

  plan(sequential)
} else if (argv$method %in% c("methyLImp2_fork", "methyLImp2_psock")) {
  library(methyLImp2)
  library(BiocParallel)
  annotation_df <- tidyr::unnest(group_df, cols = features)
  names(annotation_df) <- c("cpg", "chr")
  BPPARAM <- if (argv$cores > 1) {
    if (argv$method == "methyLImp2_psock") {
      SnowParam(workers = argv$cores, type = "SOCK", progressbar = .progress)
    } else {
      MulticoreParam(workers = argv$cores, progressbar = .progress)
    }
  } else {
    SerialParam(progressbar = .progress)
  }
  results <- methyLImp2(
    input = beta_matrix,
    type = "user",
    annotation = annotation_df,
    BPPARAM = BPPARAM
  )
}
```

## Slurm
```{r}
library(tidyverse)

all_jobs <- tidyr::expand_grid(
  method = c("knn", "knn_tree", "impute.knn", "missMDA", "pca", "methyLImp2_fork", "methyLImp2_psock"),
  cores = c(1, 2, 4, 8, 12),
  rep = 1:5
) |>
  filter(!(method == "methyLImp2_fork" & cores %in% c(1, 2, 8, 12)))

all_jobs$output <- with(all_jobs, glue::glue("output/{method}-{cores}-{rep}"))

all_jobs$job <- with(all_jobs, glue::glue(
  '/usr/bin/time -f "%U\\t%S\\t%e\\t%P\\t%M" -o {output}.txt',
  "apptainer exec $APPTAINER/missRanger/missRanger.sif Rscript benchmark.R",
  "--beta input/GSE286313.EPICv2.chr1-Y.qs2 --group input/GSE286313.EPICv2.chr1-Y.group_df.qs2",
  "--cores {cores} --method {method}",
  .sep = " "
))

all_jobs <- tidyr::nest(all_jobs, .by = "cores")

all_jobs$joblist <- here::here(
  "benchmark",
  "multi_cores",
  glue::glue("joblist_{all_jobs$cores}.txt")
)

purrr::walk2(
  all_jobs$data,
  all_jobs$joblist,
  \(x, y) {
    readr::write_lines(x$job, y)
  }
)

all_jobs$dsq <- glue::glue(
  "dsq --job-file input/joblist_{all_jobs$cores}.txt --batch-file multi_cores_{all_jobs$cores}.sh",
  "--job-name multi_cores_{all_jobs$cores} --max-jobs {c(10, 10, 10, 10, 5)}",
  "--output dSQ/dsq-multi_cores_{all_jobs$cores}-%A_%a-%N.out --status-dir .",
  "--time 03:00:00 --nodes 1 --ntasks-per-node 1 --cpus-per-task {all_jobs$cores}",
  "--mem {c(40, 40, 40, 40, 60)}G --partition day",
  .sep = " "
) |>
  as.character()

all_jobs$dsq
```

```{r}
library(here)
library(hhprf)
library(tidyverse)

job_status <- tibble(
  jl = here(
    "benchmark", "multi_cores",
    c(
      "job_5423567_status.tsv",
      "job_5423685_status.tsv",
      "job_5424266_status.tsv"
    )
  ),
  out = here(
    "benchmark", "multi_cores",
    c("joblist_4.txt", "joblist_8.txt", "joblist_12.txt")
  )
)

map2(job_status$jl, job_status$out, hhprf::retry_dsq)
```

## Viz
Ran the jobs using Slurm and /usr/time. We can use slurm sacc -j to query the actual RSS and Wall Time.

The output/ results are less correct but can be used as backup.
```{r}
library(tidyverse)

multi_cores <- tibble(
  files = list.files("benchmark/multi_cores/dSQ"),
  paths = list.files("benchmark/multi_cores/dSQ", full.names = TRUE)
)

multi_cores$output <- lapply(multi_cores$paths, read_lines)
multi_cores$output |>
  lapply(\(x) {
    grepv("full run", x)
  }) |>
  unique()

extractor <- function(x) {
  method <- if (any(str_detect(x, "knn_tree"))) {
    "knn_tree"
  } else if (any(str_detect(x, "impute.knn"))) {
    "impute.knn"
  } else if (any(str_detect(x, "knn"))) {
    "knn"
  } else if (any(str_detect(x, "pca"))) {
    "pca"
  } else if (any(str_detect(x, "missMDA"))) {
    "missMDA"
  } else if (any(str_detect(x, "methyLImp2_psock"))) {
    "methyLImp2_psock"
  } else if (any(str_detect(x, "methyLImp2_fork"))) {
    "methyLImp2_fork"
  } else {
    stop("Failed: ", x)
  }
  return(method)
}

multi_cores$method <- vapply(multi_cores$output, extractor, character(1))
multi_cores$cores <- as.numeric(
  str_extract(multi_cores$files, "(?<=multi_cores_)\\d+(?=-)")
)
multi_cores$job_id <- str_extract(multi_cores$files, "(?<=-)[0-9]+_[0-9]+(?=-)")
multi_cores[c("files", "paths", "output")] <- NULL
multi_cores |> print(n = Inf)

cmd <- glue::glue(
  "sacct -j", "{paste0(multi_cores$job_id, collapse = ',')}",
  "--format=JobID,JobName,State,Elapsed,MaxRSS,TotalCPU",
  "--parsable > multi_cores_raw.txt",
  .sep = " "
)

cmd
```

```{r}
library(scales)
multi_cores_raw <- read_delim(
  here::here("benchmark", "multi_cores", "multi_cores_raw.txt"),
  delim = "|",
) |>
  filter(State == "COMPLETED", JobName == "batch") |>
  mutate(
    job_id = str_remove(JobID, ".batch"),
    JobID = NULL,
    State = NULL,
    `...7` = NULL,
    JobName = NULL,
    max_rss_bytes = case_when(
      str_detect(MaxRSS, "K$") ~ readr::parse_number(MaxRSS) * 1024,
      str_detect(MaxRSS, "M$") ~ readr::parse_number(MaxRSS) * 1024^2,
      TRUE ~ readr::parse_number(MaxRSS)
    ),
    MaxRSS = NULL
  )

multi_cores_output <- inner_join(
  multi_cores,
  multi_cores_raw,
  by = "job_id"
)

saveRDS(
  multi_cores_output,
  here::here("benchmark", "multi_cores", "multi_cores_output.rds")
)
```

# Accuracy
## Tune
```{r tune_imp_mcar.R, eval =  F}
library(argparser)
library(impute)
library(slideimp)
library(qs2)
library(tidyr)
library(fst)
library(yardstick)
library(missMDA)
library(mirai)
library(dplyr)
library(glue)

p <- arg_parser("tune_imp")
p <- add_argument(p, "--GSE", help = "GSE accession", type = "character")
p <- add_argument(p, "--platform", help = "EPICv1/EPICv2", type = "character")
p <- add_argument(p, "--chr", help = "chromosome", type = "character")
p <- add_argument(p, "--output", help = "output file path", type = "character")
p <- add_argument(p, "--cores", help = "n cores", type = "numeric")
p <- add_argument(p, "--global_na", help = "global_na percentage", type = "numeric")
p <- add_argument(p, "--debug", help = "debug mode", flag = TRUE)

argv <- parse_args(
  p
  # ,
  # c(
  #   "--GSE", "GSE286313",
  #   "--platform", "EPICv1",
  #   "--chr", "22",
  #   "--output", "benchmark",
  #   "--cores", "7",
  #   "--global_na", 0.05
  #   # , "--debug"
  # )
)

# input
input_path <- glue(
  "{argv$GSE}/{argv$platform}/{argv$GSE}.{argv$platform}_chr{argv$chr}.qs2"
)

stopifnot(file.exists(input_path))
if (!dir.exists(argv$output)) {
  dir.create(argv$output, recursive = TRUE)
}

# knn_imp params
# knn_imp_param <- tidyr::expand_grid(
#   k = c(3, 5, 7, 10),
#   dist_pow = c(0, 2, 3, 4),
#   method = c("euclidean"),
#   post_imp = FALSE
# )
knn_imp_param <- tidyr::expand_grid(
  k = c(10, 15, 20, 25),
  dist_pow = c(0, 4, 6, 8),
  method = c("euclidean"),
  post_imp = FALSE
)

# pca_imp params. Features in columns
pca_imp_param <- tidyr::expand_grid(
  ncp = c(25, 30, 35),
  coeff.ridge = c(0.75, 1, 1.25)
)

# input
obj <- qs_read(input_path)

# inject global NA level
length_obj <- length(obj)
set.seed(1234)
dim(obj)
obj[sample.int(length_obj, size = ceiling(argv$global_na * length_obj))] <- NA
miss <- is.na(obj)
obj <- obj[(rowSums(miss) / ncol(miss)) < 0.9, ]
message("Actual global missing % is ", round((sum(is.na(obj)) * 100) / length(obj), 2))
dim(obj)

if (argv$debug) {
  obj <- obj[1:1000, ]
}
obj_t <- t(obj)
num_na <- if (argv$debug) {
  10
} else {
  10000
}

# tuning
set.seed(1234)
knn_tune_results <- tune_imp(
  obj = obj_t,
  parameters = knn_imp_param,
  .f = "knn_imp",
  rep = 14,
  num_na = num_na,
  .progress = TRUE,
  cores = argv$cores,
  colmax = 0.91
)

# setup parallel
daemons(argv$cores)

set.seed(1234)
pca_imp_tune_results <- tune_imp(
  obj = obj_t,
  parameters = pca_imp_param,
  .f = "pca_imp",
  rep = 14,
  num_na = num_na,
  .progress = TRUE,
  colmax = 0.91
)

# shutdown parallel
daemons(0)

# Save results
met_set <- metric_set(mae, rmse, rsq)

# Process knn
knn_tune_results$metrics <- lapply(
  knn_tune_results$result,
  function(x) {
    met_set(x, truth = truth, estimate = estimate)
  }
)
knn_tune_results$result <- NULL
knn_tune_results <- tidyr::unnest(knn_tune_results, cols = "metrics")
knn_tune_results <- dplyr::mutate(knn_tune_results, global_na = argv$global_na)

write_fst(
  knn_tune_results,
  file.path(
    argv$output,
    glue::glue("knn_imp_tune_{argv$GSE}_chr{argv$chr}.{argv$global_na}.fst")
  )
)

# Process pca_imp
pca_imp_tune_results$metrics <- lapply(
  pca_imp_tune_results$result,
  function(x) {
    met_set(x, truth = truth, estimate = estimate)
  }
)
pca_imp_tune_results$result <- NULL
pca_imp_tune_results <- tidyr::unnest(pca_imp_tune_results, cols = "metrics")
pca_imp_tune_results <- dplyr::mutate(pca_imp_tune_results, global_na = argv$global_na)

write_fst(
  pca_imp_tune_results,
  file.path(
    argv$output,
    glue::glue("pca_imp_tune_{argv$GSE}_chr{argv$chr}.{argv$global_na}.fst")
  )
)
```

```{bash, eval = F}
#!/bin/bash
#SBATCH --job-name=task
#SBATCH --output=./slurm-%j.out
#SBATCH --time=02:00:00
#SBATCH --nodes=1
#SBATCH --ntasks-per-node=1
#SBATCH --cpus-per-task=22
#SBATCH --mem=50G
#SBATCH --partition=day
#SBATCH --mail-type=END

# Execute the bash script within the container
apptainer exec $APPTAINER/missRanger/missRanger.sif Rscript ./tune_imp_mcar.R \
  --GSE GSE286313 \
  --platform EPICv1 \
  --chr "22" \
  --output "benchmark" \
  --cores 21 \
  --global_na 0.05
```

## Fit
Can't inject too much NA because without any cols without any NA, the imputation fails on `{methyLImp2}`

```{r fit_imp_mcar.R}
library(argparser)
library(impute)
library(slideimp)
library(qs2)
library(tidyr)
library(fst)
library(yardstick)
library(missMDA)
library(mirai)
library(dplyr)
library(glue)
library(methyLImp2)
library(BiocParallel)

p <- arg_parser("fit_imp")
p <- add_argument(p, "--GSE", help = "GSE accession", type = "character")
p <- add_argument(p, "--platform", help = "EPICv1/EPICv2/MSA", type = "character")
p <- add_argument(p, "--chr", help = "chromosome", type = "character")
p <- add_argument(p, "--output", help = "output file path", type = "character")
p <- add_argument(p, "--cores", help = "n cores", type = "numeric")
p <- add_argument(
  p, "--global_na",
  help = "global_na percentages", type = "numeric", nargs = Inf
)
p <- add_argument(p, "--debug", help = "debug mode", flag = TRUE)

argv <- parse_args(
  p
  # ,
  # c(
  #   "--GSE", "GSE286313",
  #   "--platform", "EPICv1",
  #   "--chr", "22",
  #   "--output", "benchmark",
  #   "--cores", "7",
  #   "--global_na", c(0.01)
  # )
)

input_path <- glue(
  "{argv$GSE}/{argv$platform}/{argv$GSE}.{argv$platform}_chr{argv$chr}.qs2"
)

# File and directory validation
stopifnot(file.exists(input_path))
if (!dir.exists(argv$output)) {
  dir.create(argv$output, recursive = TRUE)
}

# knn_imp params
knn_imp_param <- tibble::tibble(
  k = 20,
  dist_pow = 4,
  method = "euclidean",
  post_imp = FALSE
)

# pca_imp params
pca_imp_param <- tibble::tibble(
  ncp = 30,
  coeff.ridge = 1,
  row.w = "n_miss",
  nb.init = 5
)

methyLImp2_param <- tibble::tibble(id = 1)
methyLImp2_wrap <- function(obj, id = NULL) {
  methyLImp2::methyLImp2(
    obj,
    type = "user",
    annotation = data.frame(cpg = colnames(obj), chr = "chr22"),
    BPPARAM = BiocParallel::SerialParam()
  )
}
# input
obj <- qs_read(input_path)

daemons(argv$cores)

met_set <- metric_set(mae, rmse, rsq)
global_na_percs <- argv$global_na
seed <- 123456
num_na <- if (argv$debug) {
  10
} else {
  10000
}
rep <- if (argv$debug) {
  1
} else {
  30
}
for (perc in global_na_percs) {
  # Inject global NA level
  set.seed(seed)
  length_obj <- length(obj)
  obj_temp <- obj
  obj_temp[sample.int(length_obj, size = floor(perc * length_obj))] <- NA
  miss <- is.na(obj_temp)
  obj_temp <- obj_temp[(rowSums(miss) / ncol(miss)) < 0.9, ]
  message(
    "Actual global missing % is ",
    round((sum(is.na(obj_temp)) * 100) / length(obj_temp), 2)
  )
  dim(obj_temp)
  if (argv$debug) {
    obj_temp <- obj_temp[1:1000, ]
  }
  obj_t <- t(obj_temp)
  # knn
  set.seed(seed)
  knn_tune_results <- tune_imp(
    obj = obj_t,
    parameters = knn_imp_param,
    .f = "knn_imp",
    rep = rep,
    num_na = num_na,
    .progress = TRUE,
    cores = argv$cores,
    colmax = 0.91
  )
  knn_tune_results$metrics <- lapply(
    knn_tune_results$result,
    function(x) {
      met_set(x, truth = truth, estimate = estimate)
    }
  )
  knn_tune_results$result <- NULL
  temp_knn <- tidyr::unnest(knn_tune_results, cols = "metrics")
  temp_knn <- dplyr::mutate(temp_knn, global_na = perc)
  write_fst(
    temp_knn,
    file.path(
      argv$output,
      glue::glue("knn_imp_fitted_{argv$GSE}_{argv$platform}_chr{argv$chr}.{perc}.fst")
    )
  )
  # pca_imp
  set.seed(seed)
  pca_imp_tune_results <- tune_imp(
    obj = obj_t,
    parameters = pca_imp_param,
    .f = "pca_imp",
    rep = rep,
    num_na = num_na,
    .progress = TRUE,
    cores = argv$cores,
    colmax = 0.91
  )
  pca_imp_tune_results$metrics <- lapply(
    pca_imp_tune_results$result,
    function(x) {
      met_set(x, truth = truth, estimate = estimate)
    }
  )
  pca_imp_tune_results$result <- NULL
  temp_pca_imp <- tidyr::unnest(pca_imp_tune_results, cols = "metrics")
  temp_pca_imp <- dplyr::mutate(temp_pca_imp, global_na = perc)
  write_fst(
    temp_pca_imp,
    file.path(
      argv$output,
      glue::glue("pca_imp_fitted_{argv$GSE}_{argv$platform}_chr{argv$chr}.{perc}.fst")
    )
  )
  # methyLImp2
  set.seed(seed)
  methyLImp2_tune_results <- tune_imp(
    obj = obj_t,
    parameters = methyLImp2_param,
    .f = methyLImp2_wrap,
    rep = rep,
    num_na = num_na,
    .progress = TRUE,
    cores = argv$cores,
    colmax = 0.91
  )
  methyLImp2_tune_results$metrics <- lapply(
    methyLImp2_tune_results$result,
    function(x) {
      met_set(x, truth = truth, estimate = estimate)
    }
  )
  methyLImp2_tune_results$result <- NULL
  temp_methyLImp2 <- tidyr::unnest(methyLImp2_tune_results, cols = "metrics")
  temp_methyLImp2 <- dplyr::mutate(temp_methyLImp2, global_na = perc)
  write_fst(
    temp_methyLImp2,
    file.path(
      argv$output,
      glue::glue(
        "methyLImp2_fitted_{argv$GSE}_{argv$platform}_chr{argv$chr}.{perc}.fst"
      )
    )
  )
}

daemons(0)
```

```{r}
joblist <- glue::glue(
  "apptainer exec $APPTAINER/missRanger/missRanger.sif Rscript fit_imp_mcar.R",
  "--GSE GSE286313",
  "--platform EPICv1",
  "--chr 22",
  "--output benchmark",
  "--cores 15",
  "--global_na {c(0, 0.01, 0.025, 0.05)}",
  .sep = " "
)

# dsq_gen(
#   "joblist.txt",
#   batch_file = "fit_imp_mcar.sh",
#   job_name = "fit_imp_mcar",
#   output = "dSQ/dsq-jobfile-%A_%a-%N.out",
#   max_jobs = 2,
#   slurm_args = slurm_par(
#     time = "02:00:00",
#     cpus_per_task = "15",
#     mem = "30G"
#   )
# )

# dsq --job-file joblist.txt --batch-file fit_imp_mcar.sh --job-name fit_imp_mcar --max-jobs 2 --output dSQ/dsq-jobfile-%A_%a-%N.out --status-dir . --time 02:00:00 --nodes 1 --ntasks-per-node 1 --cpus-per-task 15 --mem 30G --partition day
```

## `group_imp`
Benchmark accuracy of `group_imp` vs full `knn_imp` and `pca_imp` on the chromosome 1:22 of the MSA chip. Focus on values in just `400` random columns each with `25` values (10,000 missing values to measure RMSE/MAE/RSQ). Do this for multiple reps.

```{r}
# Get MSA data
MSA <- get_DNAm("MSA", glue::glue("chr{c(1:22, 'X')}"))
MSA$data <- lapply(MSA$path, qs2::qs_read)
col_order <- colnames(MSA$data[[1]])
MSA$data <- lapply(MSA$data, \(x) x[, col_order])
stopifnot(length(unique(lapply(MSA$data, colnames))) == 1)

MSA <- t(purrr::reduce(MSA$data, rbind))
dim(MSA)
miss <- is.na(MSA)
MSA <- MSA[, (colSums(miss) / nrow(miss)) < 0.9]
dim(MSA)
cv <- col_vars(MSA)
MSA <- MSA[, !(is.na(cv) & dplyr::near(cv, 0))]
dim(MSA)

# Get manifest. Needed for `group_imp`
MSA_manifest <- ilmn_manifest("MSA")
all_probes <- intersect(colnames(MSA), MSA_manifest$feature_id)

# Align
MSA_manifest <- subset(MSA_manifest, feature_id %in% all_probes)
MSA <- MSA[, all_probes]

# Select columns
n_rep <- 30
set.seed(1234)

missing_2d <- tibble(
  rep = seq_len(n_rep),
  IlmnID = lapply(rep, \(x) sample(MSA_manifest$feature_id, size = 400))
) |>
  unnest(cols = "IlmnID") |>
  mutate(
    col_index = match(IlmnID, colnames(MSA)),
    row_index = lapply(IlmnID, \(x) {
      # for each column, 25 values are missing
      sample.int(nrow(MSA), size = 25)
    })
  ) |>
  unnest(cols = "row_index")

na_positions <- missing_2d |>
  nest(.by = "rep") |>
  mutate(location = lapply(
    data, \(x) {
      m <- cbind(x$row_index, x$col_index)
      colnames(m) <- c("row", "col")
      m
    }
  ))

na_positions$subset <- lapply(
  na_positions$data,
  \(x) unique(x$IlmnID)
)

na_positions$group_df <- lapply(
  na_positions$subset,
  \(x) group_features(MSA, features_df = MSA_manifest, subset = x)
)

na_positions[, c("data")] <- NULL
if (!dir.exists("benchmark/group_imp/input")) {
  dir.create("benchmark/group_imp/input")
}
qs2::qs_save(na_positions, "benchmark/group_imp/input/na_positions.qs2")
qs2::qs_save(MSA, "benchmark/group_imp/input/MSA.qs2")
```

### `group_imp.R`
```{r, group_imp.R}
library(argparser)
library(tidyverse)
library(slideimp)
library(fst)
library(glue)
library(qs2)

# Set up argument parser
p <- arg_parser("Run group/full imputation with knn/pca method")
p <- add_argument(p, "--na_positions", help = "Path to na_positions object")
p <- add_argument(p, "--MSA", help = "Path to MSA object")
p <- add_argument(p, "--output", help = "Output path")
p <- add_argument(p, "--rep", help = "Repeat number", type = "integer")
p <- add_argument(p, "--method", help = "Method: pca or knn")
p <- add_argument(p, "--group", help = "Group type: group or full")

argv <- parse_args(
  p
  # ,
  # c(
  #   "--na_positions", "benchmark/group_imp/input/na_positions.qs2",
  #   "--MSA", "benchmark/group_imp/input/MSA.qs2",
  #   "--output", "benchmark/group_imp/output",
  #   "--rep", 1,
  #   "--method", "knn",
  #   "--group", "group"
  # )
)

# Check that paths exist
stopifnot(file.exists(argv$na_position))
stopifnot(file.exists(argv$MSA))
stopifnot(dir.exists(argv$output))

# Load data
na_positions <- qs_read(argv$na_position)
MSA <- qs_read(argv$MSA)

# Output file path
results <- file.path(
  argv$output,
  glue::glue("{argv$method}-{argv$group}-{argv$rep}.fst")
)

# Set parameters
k <- 25
ncp <- 25

set.seed(12345)

# Wrapper functions
group_imp_knn_wrap <- function(obj, group_df) {
  imputed <- group_imp(
    obj = obj,
    k = k,
    group = group_df
  )[, colnames(obj)]
  return(imputed)
}

group_imp_pca_wrap <- function(obj, group_df) {
  imputed <- group_imp(
    obj = obj,
    ncp = ncp,
    group = group_df
  )[, colnames(obj)]
  return(imputed)
}

knn_imp_wrap <- function(obj, subset) {
  imputed <- knn_imp(
    obj = obj,
    k = k,
    subset = subset
  )
  obj[, subset] <- imputed[, subset]
  return(obj)
}

pca_imp_wrap <- function(obj, subset) {
  imputed <- pca_imp(
    obj = obj,
    ncp = ncp
  )[, colnames(obj)]
  return(imputed)
}

# Get the data for this specific rep
i <- argv$rep
message("Processing rep ", i)

# Get the position of missing for `tune_imp`
position <- list(na_positions$location[[i]])
# Get the column names for this rep
subset <- list(na_positions$subset[[i]])
# Get the group_df for this rep
group_df <- list(na_positions$group_df[[i]])

# Run the appropriate combination based on method and group
if (argv$method == "knn" && argv$group == "full") {
  message("Running KNN Full")
  tune_result <- tune_imp(
    obj = MSA,
    parameters = tibble::tibble(subset = subset),
    .f = knn_imp_wrap,
    rep = position
  )
  imputed <- tune_result$result[[1]]
} else if (argv$method == "knn" && argv$group == "group") {
  message("Running KNN Group")
  tune_result <- tune_imp(
    obj = MSA,
    parameters = tibble::tibble(group_df = group_df),
    .f = group_imp_knn_wrap,
    rep = position
  )
  imputed <- tune_result$result[[1]]
} else if (argv$method == "pca" && argv$group == "full") {
  message("Running PCA Full")
  tune_result <- tune_imp(
    obj = MSA,
    parameters = tibble::tibble(subset = subset),
    .f = pca_imp_wrap,
    rep = position
  )
  imputed <- tune_result$result[[1]]
} else if (argv$method == "pca" && argv$group == "group") {
  message("Running PCA Group")
  tune_result <- tune_imp(
    obj = MSA,
    parameters = tibble::tibble(group_df = group_df),
    .f = group_imp_pca_wrap,
    rep = position
  )
  imputed <- tune_result$result[[1]]
} else {
  stop("Method must be 'knn' or 'pca', group must be 'full' or 'group'")
}

# Save results
message("Saving results to ", results)
write_fst(imputed, results)
message("Done!")
```

### dSQ
```{r}
joblist_tbl <- tidyr::expand_grid(
  "na_positions" = "input/na_positions.qs2",
  "MSA" = "input/MSA.qs2",
  "output" = "output",
  "rep" = 1:30,
  "method" = c("pca", "knn"),
  "group" = c("group", "full")
)

joblist_tbl$joblist <- with(
  joblist_tbl,
  glue::glue(
    "apptainer exec $APPTAINER/missRanger/missRanger.sif Rscript group_imp.R",
    "--na_positions", "{na_positions}",
    "--MSA", "{MSA}",
    "--output", "{output}",
    "--rep", "{rep}",
    "--method", "{method}",
    "--group", "{group}",
    .sep = " "
  )
)

readr::write_lines(joblist_tbl$joblist, "benchmark/group_imp/joblist.txt")

dsq_gen(
  "joblist.txt",
  batch_file = "group_imp.sh",
  job_name = "group_imp",
  output = "dSQ/dsq-jobfile-%A_%a-%N.out",
  max_jobs = 30,
  slurm_args = slurm_par(
    time = "12:00:00",
    cpus_per_task = "1",
    mem = "30G"
  )
)

# dsq --job-file joblist.txt --batch-file group_imp.sh --job-name group_imp --max-jobs 30 --output dSQ/dsq-jobfile-%A_%a-%N.out --status-dir . --time 12:00:00 --nodes 1 --ntasks-per-node 1 --cpus-per-task 1 --mem 30G --partition day
```

### Viz
```{r}
group_imp <- tibble(
  files = list.files("benchmark/group_imp/output"),
  paths = list.files("benchmark/group_imp/output", full.names = TRUE)
) |>
  separate_wider_delim(
    cols = "files",
    delim = "-", names = c("method", "group", "rep")
  ) |>
  mutate(result = lapply(paths, fst::read_fst))

class(group_imp) <- c("TuneImp", "tbl_df", "tbl", "data.frame")

group_imp <- compute_metrics(group_imp, c("mae", "rmse", "rsq", "bias"))
group_imp$paths <- NULL
group_imp$rep <- as.numeric(str_remove(group_imp$rep, ".fst"))

fst::write_fst(group_imp, "benchmark/full_vs_group.fst")
```

